### 2016-09-15 revision/ada-ftrl.tex
new directory for revision and first draft

In this section we generalize the \textsc{AdaHedge}
algorithm~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to the OLO setting,
showing that it retains its scale-free property. The analysis is very general
and based on general properties of strongly convex functions, rather than
specific properties of the entropic regularizer like in \textsc{AdaHedge}.

@@

The regret bound can be optimized by choosing the optimal multiple of the
regularizer.  Namely, we choose regularizer of the form $\lambda f(w)$ where
$f(w)$ is $1$-strongly convex and optimize over $\lambda$. The result of the
optimization is the following corollary.  Its proof can be found
in~\ref{section:ada-ftrl-proofs}.

\begin{corollary}[Regret Bound]
\label{corollary:ada-ftrl-regret-bound}
Suppose $K \subseteq V$ is a non-empty bounded closed convex set. Suppose $f:K
\to \R$ is a non-negative lower semi-continuous function that is $1$-strongly
convex with respect to $\|\cdot\|$ and is bounded from above.  The regret of
\textsc{AdaFTRL} with regularizer
$$
R(w) = \frac{f(w)}{16 \cdot \sup_{v \in K} f(v)}
\qquad \text{satisfies} \qquad
\Regret_T \le
5.3 \sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T \|\ell_t\|_*^2} \; .
$$
\end{corollary}

@@

\begin{lemma}[Solution of the Recurrence]
\label{lemma:recurrence-solution}
Let $\{a_t\}_{t=1}^\infty$ be any sequence of non-negative real numbers.
Suppose that $\{\Delta_t\}_{t=0}^\infty$ is a sequence of real numbers
satisfying
$$
\Delta_0 = 0 \qquad \text{and} \qquad
\Delta_t \le \Delta_{t-1} + \min \left\{ a_t, \ \frac{a_t^2}{\Delta_{t-1}} \right\} \quad \text{for any $t \ge 1$} \; .
$$
Then, for any $T \ge 0$,
$$
\Delta_T \le \sqrt{3 \sum_{t=1}^T a_t^2} \; .
$$
\end{lemma}
%
Proof of the Lemma~\ref{lemma:recurrence-solution} is deferred
to~\ref{section:ada-ftrl-proofs}.  Theorem~\ref{theorem:ada-ftrl-regret-bound}
follows from
Lemmas~\ref{lemma:initial-regret-bound},~\ref{lemma:gap-recurrence}~and~\ref{lemma:recurrence-solution}.

>>>>

Here we show that it is possible to derive scale-free algorithms in an alternative way. We generalize the \textsc{AdaHedge}
algorithm~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to the OLO setting,
showing that it retains its scale-free property. The analysis is very general
and based on general properties of strongly convex functions, rather than
specific properties of the entropic regularizer like in \textsc{AdaHedge}.

@@

The regret bound can be optimized by choosing the optimal multiple of the
regularizer.  Namely, we choose regularizer of the form $\lambda f(w)$ where
$f(w)$ is $1$-strongly convex and optimize over $\lambda$. The result of the
optimization is the following corollary.

\begin{corollary}[Regret Bound]
\label{corollary:ada-ftrl-regret-bound}
Suppose $K \subseteq V$ is a non-empty bounded closed convex set. Suppose $f:K
\to \R$ is a non-negative lower semi-continuous function that is $1$-strongly
convex with respect to $\|\cdot\|$ and is bounded from above.  The regret of
\textsc{AdaFTRL} with regularizer
$$
R(w) = \frac{f(w)}{16 \cdot \sup_{v \in K} f(v)}
\qquad \text{satisfies} \qquad
\Regret_T \le
5.3 \sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T \|\ell_t\|_*^2} \; .
$$
\end{corollary}
%
\begin{proof}
Let $S = \sup_{v \in K} f(v)$. Theorem~\ref{theorem:ada-ftrl-regret-bound}
applied to the regularizer $R(w) = \frac{c}{S} f(w)$ and
Proposition~\ref{proposition:diameter-vs-range} gives
$$
\Regret_T \le \sqrt{3}(1 + c) \max\left\{\sqrt{8}, \frac{1}{\sqrt{2c}} \right\} \sqrt{S \sum_{t=1}^T \|\ell_t\|_*^2} \; .
$$
It remains to find the minimum of $g(c) = \sqrt{3}(1 + c) \max\{\sqrt{8},
1/\sqrt{2c}\}$.  The function $g$ is strictly convex on $(0, \infty)$ and has
minimum at $c=1/16$ and $g(\frac{1}{16}) = \sqrt{3}(1+\frac{1}{16})\sqrt{8} \le
5.3$.
\end{proof}

@@

\begin{lemma}[Solution of the Recurrence]
\label{lemma:recurrence-solution}
Let $\{a_t\}_{t=1}^\infty$ be any sequence of non-negative real numbers.
Suppose that $\{\Delta_t\}_{t=0}^\infty$ is a sequence of real numbers
satisfying
$$
\Delta_0 = 0 \qquad \text{and} \qquad
\Delta_t \le \Delta_{t-1} + \min \left\{ a_t, \ \frac{a_t^2}{\Delta_{t-1}} \right\} \quad \text{for any $t \ge 1$} \; .
$$
Then, for any $T \ge 0$,
$$
\Delta_T \le \sqrt{3 \sum_{t=1}^T a_t^2} \; .
$$
\end{lemma}
%
\begin{proof}[Proof of Lemma~\ref{lemma:recurrence-solution}]
We prove the lemma by induction on $T$.  The base case $T=0$ is trivial.
For $T \ge 1$, let $f(x) = x + \min\{a_T, a_T^2/x\}$. We have
\begin{align*}
\Delta_T
& \le \Delta_{T-1} + \min \left\{a_T, \ \frac{a_T^2}{\Delta_{T-1}} \right\} \\
& = f \left( \Delta_{T-1} \right) \\
& \le f \left( \sqrt{3 \sum_{t=1}^{T-1} a_t^2} \right) \\
& = \sqrt{3 \sum_{t=1}^{T-1} a_t^2} \ + \ \min \left\{ a_T, \ \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}  \; ,
\end{align*}
where the second inequality follows from the inductive hypothesis $\Delta_{T-1} \le
\sqrt{3 \sum_{t=1}^{T-1} a_t^2}$ and that $f(x)$ is non-positive on $[-\infty,
0)$ and non-negative increasing on $[0,\infty)$. It remains to prove that
$$
\sqrt{3 \sum_{t=1}^{T-1} a_t^2} \ + \ \min \left\{ a_T, \ \frac{a_T^2}{\sqrt{3 \sum_{t=1}^{T-1} a_t^2}} \right\}
\le  \sqrt{3 \sum_{t=1}^T a_t^2} \; .
$$
If $a_T = 0$, the inequality clearly holds. Otherwise, $a_T$ is positive and we
divide through by it and make substitution $z=\frac{\sqrt{\sum_{t=1}^{T-1}
a_t^2}}{a_T}$, which leads to
$$
z\sqrt{3} + \min\left\{1, \frac{1}{z\sqrt{3}}\right\} \le \sqrt{3 + 3z^2}
$$
which can be easily checked by considering separately the cases $z \in
[0,\frac{1}{\sqrt{3}})$ and $z \in [\frac{1}{\sqrt{3}}, \infty)$.
\end{proof}

Theorem~\ref{theorem:ada-ftrl-regret-bound}
follows from
Lemmas~\ref{lemma:initial-regret-bound},~\ref{lemma:gap-recurrence}~and~\ref{lemma:recurrence-solution}.

### 2016-09-15 revision/introduction.tex
new directory for revision and first draft

For this reason, most of the existing OLO algorithms receive as an input---or
explicitly assume---an upper bound $B$ on the norm of the loss vectors.  The
input $B$ is often disguised as the learning rate, the regularization
parameter, or the parameter of strong convexity of the regularizer. Examples of
such algorithms include \textsc{Hedge}~\cite{Freund-Schapire-1997} or online
projected gradient descent with fixed learning rate.  However, these algorithms
have two obvious drawbacks.

@@

\begin{table}[t]
\fontsize{8}{8.2}\selectfont
\centering
\begin{tabular}{|p{3.6cm}|c|p{3.4cm}|c|}
\hline
\textbf{Algorithm} & \textbf{Decisions Set(s)} & \textbf{Regularizer(s)} & \textbf{Scale-Free} \\ \hline \hline
\textsc{Hedge} \cite{Freund-Schapire-1997} & Probability Simplex & Negative Entropy & No \\ \hline
\textsc{GIGA} \cite{Zinkevich-2003} & Any Bounded & $\frac{1}{2}\|w\|_2^2$ & No \\ \hline
\textsc{RDA} \cite{Xiao-2010} & \textbf{Any} & \textbf{Any Strongly Convex} & No \\ \hline
\textsc{FTRL-Proximal} \cite{McMahan-Streeter-2010,McMahan-2014} & Any Bounded & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & \textbf{Yes} \\ \hline
\textsc{AdaGrad MD} \cite{Duchi-Hazan-Singer-2011} & Any Bounded & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & \textbf{Yes} \\ \hline
\textsc{AdaGrad FTRL} \cite{Duchi-Hazan-Singer-2011} & \textbf{Any} & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & No \\ \hline
\textsc{AdaHedge} \cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} & Probability Simplex & Negative Entropy & \textbf{Yes} \\ \hline
\textsc{Optimistic MD} \cite{Rakhlin-Sridharan-2013} & $\sup_{u,v \in K} \Breg_f(u,v) < \infty$ & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\textsc{NAG} \cite{Ross-Mineiro-Langford-2013} & $\{u: \max_t \langle \ell_t, u\rangle \le C\}$ & $\frac{1}{2}\|w\|_2^2 $& Partially\footnotemark \\ \hline
\textsc{Scale invariant algorithms} \cite{Orabona-Crammer-Cesa-Bianchi-2014} & \textbf{Any} & $\frac{1}{2}\|w\|_p^2 + $ any convex func. \newline $1 < p \le 2$ & Partially\textsuperscript{\ref{footnote-label}} \\ \hline
\textsc{AdaFTRL} \textbf{[this paper]} & Any Bounded & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\textsc{SOLO FTRL} \textbf{[this paper]} & \textbf{Any} & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\end{tabular}
\caption{Selected results for OLO. Best results in each column are in bold.
\label{table:results}
}
\end{table}

\footnotetext{\label{footnote-label} These algorithms attempt to produce an
invariant sequence of predictions $\langle w_t, \ell_t \rangle$, rather than a
sequence of invariant $w_t$.}

@@

Majority of existing algorithms for OLO are based on two generic algorithms:
\textsc{Follow The Regularizer Leader} (\textsc{FTRL}) and \textsc{Mirror
Descent} (\textsc{MD}). \textsc{FTRL} dates back to the potential-based
forecaster in \cite[Chapter~11]{Cesa-Bianchi-Lugosi-2006} and its theory was
developed in \cite{Shalev-Shwartz-2007}.  The name \textsc{Follow The
Regularized Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}.
Independently, the same algorithm was proposed in~\cite{Nestorov-2009} for
convex optimization under the name \textsc{Dual Averaging} and later
rediscovered in~\cite{Xiao-2010} for online convex optimization. The general
version with time-varying regularizers was analyzed in
\cite{Orabona-Crammer-Cesa-Bianchi-2014}. \textsc{MD} was originally proposed
in \cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003}
for convex optimization. In the online learning literature it makes the first
appearance, under a different name, in \cite{Kivinen-Warmuth-1997}. A good
description of \textsc{MD} is given in~\cite{Rakhlin-Sridharan-2009}.

Both \textsc{FTRL} and \textsc{MD} are parametrized by a function called a
\emph{regularizer}. The function is assumed to be strongly convex with respect
to a given norm. Based on different regularizers different algorithms with
different properties can be instantiated. A summary of algorithms for OLO is
presented in Table~\ref{table:results}.  All of them are instances of
\textsc{FTRL} or \textsc{MD}.

Scale-free versions of \textsc{MD} include \textsc{AdaGrad
MD}~\cite{Duchi-Hazan-Singer-2011} and~\textsc{Optimistic
MD}~\cite{Rakhlin-Sridharan-2013}. However, these algorithms have non-trivial
regret bounds only when the Bregman divergence associated with the regularizer
is bounded. In particular, since a bound on the Bregman divergence implies that
the decision set is bounded, the regret bounds for \textsc{AdaGrad MD} and
\textsc{Optimistic MD} are vacuous for unbounded sets. In fact, as we show in
Section~\ref{subsection:mirror-descent-lower-bound}, \textsc{AdaGrad MD} and
similar algorithms based on \textsc{MD} have $\Omega(T)$ regret if the Bregman
divergence is not bounded.

Only one scale-free algorithm based on \textsc{FTRL} was known. It is the
\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} algorithm for
learning with expert advice, where the decision set is bounded.
Algorithms based on \textsc{FTRL} that are ``almost'' scale-free include
\textsc{FTRL-Proximal}~\cite{McMahan-Streeter-2010,McMahan-2014} and
\textsc{AdaGrad FTRL}~\cite{Duchi-Hazan-Singer-2011}.  These algorithms fail to
be scale-free due to ``off-by-one'' issue; see~\cite{McMahan-2014} and also our
discussion in Section~\ref{section:solo-ftrl}.

@@

We design and analyze three scale-free algorithms: \textsc{AdaFTRL},
\textsc{SOLO FTRL} and \textsc{Scale-Free MD}.  \textsc{AdaFTRL} and
\textsc{SOLO FTRL} are based on \textsc{FTRL}.  \textsc{AdaFTRL} is a
generalization of
\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to arbitrary
strongly convex regularizers.  \textsc{SOLO FTRL} can be viewed as the
``correct'' scale-free version of \textsc{AdaGrad
FTRL}~\cite{Duchi-Hazan-Singer-2011} generalized to arbitrary strongly convex
regularizers.  \textsc{Scale-Free MD} is based on \textsc{MD}. It is a
generalization of \textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011} to
arbitrary strongly convex regularizers, similar to \textsc{Optimistic
MD}~\cite{Rakhlin-Sridharan-2013}.  The three algorithms are presented and
analyzed in Sections~\ref{section:ada-ftrl}, \ref{section:solo-ftrl} and
\ref{section:mirror-descent} respectively.

We prove that the regret of \textsc{AdaFTRL} and \textsc{SOLO FTRL} after $T$
rounds is bounded by $O \left(\sqrt{\sup_{v \in K} f(v)
\sum_{t=1}^T\|\ell_t\|_*^2} \right)$ where $f$ is a non-negative regularizer
that is $1$-strongly convex with respect to a norm $\|\cdot\|$ and
$\|\cdot\|_*$ is its dual norm. For \textsc{Scale-Free MD}, we prove $O
\left(\sqrt{\sup_{u,v \in K} B_f(u,v) \sum_{t=1}^T\|\ell_t\|_*^2} \right)$
where $B_f$ is the Bregman divergence associated with a $1$-strongly convex
regularizer $f$.  As we discuss, in Section~\ref{section:mirror-descent}, the
upper bound for \textsc{Scale-Free MD} is never better than the bound for bound
for \textsc{AdaFTRL} and \textsc{SOLO FTRL}.  In
Section~\ref{section:lower-bound}, we show that the $\sqrt{\sum_{t=1}^T
\|\ell_t\|_*^2}$ term in the bounds is necessary by proving $\frac{D}{\sqrt{8}}
\sqrt{\sum_{t=1}^T\|\ell_t\|_*^2}$ lower bound on the regret of any algorithm
for OLO for any decision set with diameter $D$ with respect to the primal norm
$\|\cdot\|$.

For \textsc{SOLO FTRL}, we prove that regret against a competitor $u \in K$ is
at most $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
\|\ell_t\|_* \sqrt{T} \right)$.  Same as before, $f$ is a non-negative
$1$-strongly convex regularizer. This bound is non-trivial for any decision
set, bounded or unbounded.  The result makes \textsc{SOLO FTRL} the
\textbf{first adaptive algorithm for unbounded decision sets} with a
non-trivial regret bound.

@@

Finally, in Section~\ref{subsection:mirror-descent-lower-bound}, we show
negative results for existing popular variants of \textsc{MD}. We show two
examples that demonstrate that variants of \textsc{MD} can have $\Omega(T)$
regret (on sequences of loss vectors of unit norm) if the Bregman divergence
associated with the regularizer is unbounded.  First example demonstrates
$\Omega(T^{3/2})$ regret of the standard gradient descent with step size
$1/\sqrt{t}$, \textsc{AdaGrad MD} and \textsc{Scale-Free MD} on a unbounded
decision set.  Second example demonstrates $\Omega(T)$ regret for
\textsc{Scale-Free MD} for learning with expert advice with negative entropy
regularizer.  Both results indicate that \textsc{FTRL} is superior to
\textsc{MD}.

>>>>

For this reason, most of the existing OLO algorithms receive as an input---or
explicitly assume---an upper bound $B$ on the norm of the loss vectors.  The
input $B$ is often disguised as the learning rate, the regularization
parameter, or the parameter of strong convexity of the regularizer.
%Examples of
%such algorithms include \textsc{Hedge}~\cite{Freund-Schapire-1997} or online
%projected gradient descent with fixed learning rate.
However, these algorithms have two obvious drawbacks.

@@

\begin{table}[t]
\fontsize{8}{8.2}\selectfont
\centering
\begin{tabular}{|p{3.6cm}|c|p{3.4cm}|c|}
\hline
\textbf{Algorithm} & \textbf{Decisions Set(s)} & \textbf{Regularizer(s)} & \textbf{Scale-Free} \\ \hline \hline
\textsc{Hedge} \cite{Freund-Schapire-1997} & Probability Simplex & Negative Entropy & No \\ \hline
\textsc{GIGA} \cite{Zinkevich-2003} & Any Bounded & $\frac{1}{2}\|w\|_2^2$ & No \\ \hline
\textsc{RDA} \cite{Xiao-2010} & \textbf{Any} & \textbf{Any Strongly Convex} & No \\ \hline
\textsc{FTRL-Proximal} \cite{McMahan-Streeter-2010,McMahan-2014} & Any Bounded & $\frac{1}{2}\|w\|_2^2 + $ any convex func.\footnotemark & \textbf{Yes} \\ \hline
\textsc{AdaGrad MD} \cite{Duchi-Hazan-Singer-2011} & Any Bounded & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & \textbf{Yes} \\ \hline
\textsc{AdaGrad FTRL} \cite{Duchi-Hazan-Singer-2011} & \textbf{Any} & $\frac{1}{2}\|w\|_2^2 + $ any convex func. & No \\ \hline
\textsc{AdaHedge} \cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} & Probability Simplex & Negative Entropy & \textbf{Yes} \\ \hline
\textsc{NAG} \cite{Ross-Mineiro-Langford-2013} & $\{u: \max_t \langle \ell_t, u\rangle \le C\}$ & $\frac{1}{2}\|w\|_2^2 $& Partially\footnotemark \\ \hline
\textsc{Scale invariant algorithms} \cite{Orabona-Crammer-Cesa-Bianchi-2014} & \textbf{Any} & $\frac{1}{2}\|w\|_p^2 + $ any convex func. \newline $1 < p \le 2$ & Partially\textsuperscript{\ref{footnote-label2}} \\ \hline
%\textsc{AdaFTRL} \textbf{[this paper]} & Any Bounded & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\textsc{Scale-free MD} \textbf{[folklore]} & $\sup_{u,v \in K} \Breg_f(u,v) < \infty$ & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\textsc{SOLO FTRL} \textbf{[this paper]} & \textbf{Any} & \textbf{Any Strongly Convex} & \textbf{Yes} \\ \hline
\end{tabular}
\caption{Selected results for OLO. Best results in each column are in bold.
\label{table:results}
}
\end{table}

\addtocounter{footnote}{-2}

\stepcounter{footnote}\footnotetext{\label{footnote-label1} Even if, in principle the FTRL-Proximal algorithm can be used with any proximal regularizer, to the best of our knowledge a general way to construct proximal regularizers is not known. The only proximal regularizer we are aware is based on the 2-norm.}

\stepcounter{footnote}\footnotetext{\label{footnote-label2} These algorithms attempt to produce an
invariant sequence of predictions $\langle w_t, \ell_t \rangle$, rather than a
sequence of invariant $w_t$.}

@@

The majority of the existing algorithms for OLO are based on two generic algorithms:
\textsc{Follow The Regularizer Leader} (\textsc{FTRL}) and \textsc{Mirror
Descent} (\textsc{MD}). \textsc{FTRL} dates back to the potential-based
forecaster in \cite[Chapter~11]{Cesa-Bianchi-Lugosi-2006} and its theory was
developed in \cite{Shalev-Shwartz-2007}.  The name \textsc{Follow The
Regularized Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}.
Independently, the same algorithm was proposed in~\cite{Nestorov-2009} for
convex optimization under the name \textsc{Dual Averaging} and rediscovered
in~\cite{Xiao-2010} for online convex optimization. The general
version with time-varying regularizers was analyzed in
\cite{Orabona-Crammer-Cesa-Bianchi-2014}. \textsc{MD} was originally proposed
in \cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003}
for convex optimization. In the online learning literature it makes his first
appearance, with a different name, in~\cite{Kivinen-Warmuth-1997}.
%A good description of \textsc{MD} is given in~\cite{Rakhlin-Sridharan-2009}.

Both \textsc{FTRL} and \textsc{MD} are parametrized by a function called 
\emph{regularizer}. The function is assumed to be strongly convex with respect
to a given norm. Based on different regularizers different algorithms with
different properties can be instantiated. A summary of algorithms for OLO is
presented in Table~\ref{table:results}.  All of them are instances of
\textsc{FTRL} or \textsc{MD}.

Scale-free versions of \textsc{MD} include \textsc{AdaGrad
MD}~\cite{Duchi-Hazan-Singer-2011}. However, the AdaGrad
MD algorithm has a non-trivial regret bounds only when the Bregman divergence associated with the regularizer
is bounded. In particular, since a bound on the Bregman divergence implies that
the decision set is bounded, the regret bound for \textsc{AdaGrad MD} and
is vacuous for unbounded sets. In fact, as we show in
Section~\ref{subsection:mirror-descent-lower-bound}, \textsc{AdaGrad MD} and
similar algorithms based on \textsc{MD} have $\Omega(T)$ regret if the Bregman
divergence is not bounded.

Only one scale-free algorithm based on \textsc{FTRL} was known. It is the
\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} algorithm for
learning with expert advice, where the decision set is bounded.
An algorithm based on \textsc{FTRL} that is ``almost'' scale-free is 
\textsc{AdaGrad FTRL}~\cite{Duchi-Hazan-Singer-2011}.  This algorithm fail to
be scale-free due to ``off-by-one'' issue; see~\cite{McMahan-2014} and the
discussion in Section~\ref{section:solo-ftrl}. Instead, \textsc{FTRL-Proximal}~\cite{McMahan-Streeter-2010,McMahan-2014} solves the off-by-one issues, but it requires proximal regularizers, that in general do not have a simple form and even in the simple 2-norm case requires bounded domains.

@@

We design and analyze three scale-free algorithms: \textsc{AdaFTRL},
\textsc{SOLO FTRL} and \textsc{Scale-Free MD}.  \textsc{AdaFTRL} and
\textsc{SOLO FTRL} are based on \textsc{FTRL}.  \textsc{AdaFTRL} is a
generalization of
\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to arbitrary
strongly convex regularizers.  \textsc{SOLO FTRL} can be viewed as the
``correct'' scale-free version of the diagonal version of \textsc{AdaGrad
FTRL}~\cite{Duchi-Hazan-Singer-2011} generalized to arbitrary strongly convex
regularizers.  \textsc{Scale-Free MD} is based on \textsc{MD}. It is a
generalization of \textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011} to
arbitrary strongly convex regularizers.  The three algorithms are presented in
Sections~\ref{section:ada-ftrl}, \ref{section:solo-ftrl} and
\ref{section:mirror-descent} respectively.

We prove that the regret of \textsc{AdaFTRL} and \textsc{SOLO FTRL} after $T$
rounds is bounded by $O (\sqrt{\sup_{v \in K} f(v)
\sum_{t=1}^T\|\ell_t\|_*^2} )$ where $f$ is a non-negative regularizer
that is $1$-strongly convex with respect to a norm $\|\cdot\|$ and
$\|\cdot\|_*$ is its dual norm. For \textsc{Scale-Free MD}, we prove $O
(\sqrt{\sup_{u,v \in K} B_f(u,v) \sum_{t=1}^T\|\ell_t\|_*^2} )$
where $B_f$ is the Bregman divergence associated with a $1$-strongly convex
regularizer $f$.  As we discuss, in Section~\ref{section:mirror-descent}, the
upper bound for \textsc{Scale-Free MD} is never better than the bound for bound
for \textsc{AdaFTRL} and \textsc{SOLO FTRL}.  In
Section~\ref{section:lower-bound}, we show that the $\sqrt{\sum_{t=1}^T
\|\ell_t\|_*^2}$ term in the bounds is necessary by proving $\frac{D}{\sqrt{8}}
\sqrt{\sum_{t=1}^T\|\ell_t\|_*^2}$ lower bound on the regret of any algorithm
for OLO for any decision set with diameter $D$ with respect to the primal norm
$\|\cdot\|$.

For \textsc{SOLO FTRL}, we prove that regret against a competitor $u \in K$ is
at most $O (f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
\|\ell_t\|_* \sqrt{T} )$.  Same as before, $f$ is a non-negative
$1$-strongly convex regularizer. This bound is non-trivial for any decision
set, bounded or unbounded.  The result makes \textsc{SOLO FTRL} the
\textbf{first adaptive algorithm for unbounded decision sets} with a
non-trivial regret bound.

@@

Finally, in Section~\ref{subsection:mirror-descent-lower-bound}, we show
negative results for existing popular variants of \textsc{MD}. We show two
examples that demonstrate that variants of \textsc{MD} can have $\Omega(T)$
regret (on sequences of loss vectors of unit norm) if the Bregman divergence
associated with the regularizer is unbounded.
%First example demonstrates
%$\Omega(T^{3/2})$ regret of the standard gradient descent with step size
%$1/\sqrt{t}$, \textsc{AdaGrad MD} and \textsc{Scale-Free MD} on a unbounded
%decision set.  Second example demonstrates $\Omega(T)$ regret for
%\textsc{Scale-Free MD} for learning with expert advice with negative entropy
%regularizer.
These results indicate that \textsc{FTRL} is superior to \textsc{MD}.

### 2016-09-15 revision/lower-bound.tex
new directory for revision and first draft

We show a lower bound on the worst-case regret of any algorithm for OLO. The
proof, which we present in~\ref{section:lower-bound-proof}, is a standard
probabilistic argument.

@@

The upper bounds on regret of \textsc{AdaFTRL} and \textsc{SOLO FTRL} are of
the form $O(\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T \|\ell_t\|_*^2})$ where $f$
is any $1$-strongly convex function with respect to $\|\cdot\|$.  The same
upper bound is also achieved by \textsc{FTRL} with a constant learning rate
when the number of rounds $T$ and $\sum_{t=1}^T \|\ell_t\|_*^2$ is known
upfront \cite[Chapter 2]{Shalev-Shwartz-2011}.  The lower bound is
$\Omega(D\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$.

>>>>

We show a lower bound on the worst-case regret of any algorithm for OLO. The
proof, presented in~\ref{section:lower-bound-proof}, is a standard
probabilistic argument.

@@

The upper bound on regret of \textsc{SOLO FTRL} is of
the form $O(\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T \|\ell_t\|_*^2})$ where $f$
is any $1$-strongly convex function with respect to $\|\cdot\|$.  The same
upper bound is also achieved by \textsc{FTRL} with a constant learning rate
when the number of rounds $T$ and $\sum_{t=1}^T \|\ell_t\|_*^2$ is known
upfront \cite[Chapter 2]{Shalev-Shwartz-2011}.  The lower bound is
$\Omega(D\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$.
